---
title: "Models Update"
author: "Kyle W. Brown"
date: "4/17/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Linear Regressions

##John Hopkins University U.S. Counties COVID-19 

```{r}
set.seed(8)
countyUS <- read.csv("https://raw.githubusercontent.com/WorldCapital/COVID19-Project/master/COVID-19/JHU_county_time_series.csv")
countyUS <- subset(countyUS, select = -c(X))
str(countyUS)
head(countyUS)
```


##Confirmed Model

```{r}
confirmedUSlm <- lm(Confirmed ~ Deaths + Recovered + FIPS + Incident_Rate + People_Tested + People_Hospitalized + Mortality_Rate + Testing_Rate + Hospitalization_Rate, data = countyUS)
confint(confirmedUSlm)
summary(confirmedUSlm)

mse = mean(confirmedUSlm$residuals^2)
print(paste0("MSE= ", mse))


# Function for Root Mean Squared Error
RMSE <- function(error) { sqrt(mean(error^2)) }
print(paste0("RMSE= ", RMSE(confirmedUSlm$residuals)))


par(mfrow = c(2,2)) 
plot(confirmedUSlm)
```

## Deaths Model

```{r}
deathsUSlm <- lm(Deaths ~ Confirmed + Recovered + FIPS + Incident_Rate + People_Tested + People_Hospitalized + Mortality_Rate + Testing_Rate + Hospitalization_Rate, data = countyUS)
confint(deathsUSlm)
summary(deathsUSlm)

mse = mean(deathsUSlm$residuals^2)
print(paste0("MSE= ", mse))
print(paste0("RMSE= ", RMSE(deathsUSlm$residuals)))


par(mfrow = c(2,2)) 
plot(deathsUSlm)
```


## Recovered Model

```{r}
recoveredUSlm <- lm(Recovered ~ Confirmed + Deaths + FIPS + Incident_Rate + People_Tested + People_Hospitalized + Mortality_Rate + Testing_Rate + Hospitalization_Rate, data = countyUS)
confint(recoveredUSlm)
summary(recoveredUSlm)

mse = mean(recoveredUSlm$residuals^2)
print(paste0("MSE= ", mse))
print(paste0("RMSE= ", RMSE(recoveredUSlm$residuals)))

par(mfrow = c(2,2)) 
plot(recoveredUSlm)
```


##FIPS Model

```{r}
fipsUSlm <- lm(FIPS ~ Confirmed + Deaths + Recovered + Incident_Rate + People_Tested + People_Hospitalized + Mortality_Rate + Testing_Rate + Hospitalization_Rate, data = countyUS)
confint(fipsUSlm)
summary(fipsUSlm)

mse = mean(fipsUSlm$residuals^2)
print(paste0("MSE= ", mse))
print(paste0("RMSE= ", RMSE(fipsUSlm$residuals)))


par(mfrow = c(2,2)) 
plot(fipsUSlm)
```

## Incident_Rate Model

```{r}
incidentUSlm <- lm(Incident_Rate ~ Confirmed + Deaths + Recovered + FIPS + People_Tested + People_Hospitalized + Mortality_Rate + Testing_Rate + Hospitalization_Rate, data = countyUS)
confint(incidentUSlm)
summary(incidentUSlm)

mse = mean(incidentUSlm$residuals^2)
print(paste0("MSE= ", mse))
print(paste0("RMSE= ", RMSE(incidentUSlm$residuals)))


par(mfrow = c(2,2)) 
plot(incidentUSlm)
```


## People_Tested Model 

```{r}
testedUSlm <- lm(People_Tested ~ Confirmed + Deaths + Recovered + FIPS + Incident_Rate  + People_Hospitalized + Mortality_Rate + Testing_Rate + Hospitalization_Rate, data = countyUS)
confint(testedUSlm)
summary(testedUSlm)

mse = mean(testedUSlm$residuals^2)
print(paste0("MSE= ", mse))
print(paste0("RMSE= ", RMSE(testedUSlm$residuals)))

par(mfrow = c(2,2)) 
plot(testedUSlm)
```


## People_Hospitalized Model

```{r}
hospUSlm <- lm(People_Hospitalized ~ Confirmed + Deaths + Recovered + FIPS + Incident_Rate  + People_Tested  + Mortality_Rate + Testing_Rate + Hospitalization_Rate, data = countyUS)
confint(hospUSlm)
summary(hospUSlm)

mse = mean(hospUSlm$residuals^2)
print(paste0("MSE= ", mse))
print(paste0("RMSE= ", RMSE(hospUSlm$residuals)))


par(mfrow = c(2,2)) 
plot(hospUSlm)
```



## Mortal_Rate Model

```{r}
mortalUSlm <- lm(Mortality_Rate ~ Confirmed + Deaths + Recovered + FIPS + Incident_Rate  + People_Tested  + People_Hospitalized  + Testing_Rate + Hospitalization_Rate, data = countyUS)
confint(mortalUSlm)
summary(mortalUSlm)

mse = mean(mortalUSlm$residuals^2)
print(paste0("MSE= ", mse))
print(paste0("RMSE= ", RMSE(mortalUSlm$residuals)))


par(mfrow = c(2,2)) 
plot(mortalUSlm)
```


## Testing_Rate Model

```{r}
testRtUSlm <- lm(Testing_Rate ~ Confirmed + Deaths + Recovered + FIPS + Incident_Rate  + People_Tested  + People_Hospitalized + Mortality_Rate + Hospitalization_Rate, data = countyUS)
confint(testRtUSlm)
summary(testRtUSlm)

mse = mean(testRtUSlm$residuals^2)
print(paste0("MSE= ", mse))
print(paste0("RMSE= ", RMSE(testRtUSlm$residuals)))

par(mfrow = c(2,2)) 
plot(testRtUSlm)
```



## Hospitalization_Rate Model

```{r}
hospRtUSlm <- lm(Hospitalization_Rate ~ Confirmed + Deaths + Recovered + FIPS + Incident_Rate + People_Tested + People_Hospitalized + Mortality_Rate + Testing_Rate, data = countyUS)
confint(hospRtUSlm)
summary(hospRtUSlm)

mse = mean(hospRtUSlm$residuals^2)
print(paste0("MSE= ", mse))
print(paste0("RMSE= ", RMSE(hospRtUSlm$residuals)))


par(mfrow = c(2,2)) 
plot(hospRtUSlm)
```


# Linear Regression with COVID-19 STATWORX API

```{r}
library(httr)
library(jsonlite) 

payload <- list(code = "US") 
response <- httr::POST(url = "https://api.statworx.com/covid", body = toJSON(payload, auto_unbox = TRUE), encode = "json") 

content <- rawToChar(response$content) 
us <- data.frame(fromJSON(content)) 
str(us) 
head(us)
```

## Cumulative Cases Model

```{r}
confirmSTATlm <- lm(cases_cum ~ day + month + deaths_cum, data = us)
confint(confirmSTATlm)
summary(confirmSTATlm)

mse = mean(confirmSTATlm$residuals^2)
print(paste0("MSE= ", mse))
print(paste0("RMSE= ", RMSE(confirmSTATlm$residuals)))


par(mfrow = c(2,2)) 
plot(confirmSTATlm)
```


## Cumulative Deaths Model

```{r}
deathsSTATlm <- lm(deaths_cum ~ day + month + cases_cum, data = us)
confint(deathsSTATlm)
summary(confirmSTATlm)

mse = mean(deathsSTATlm$residuals^2)
print(paste0("MSE= ", mse))
print(paste0("RMSE= ", RMSE(deathsSTATlm$residuals)))


par(mfrow = c(2,2)) 
plot(deathsSTATlm)
```


## Day Model

```{r}
daySTATlm <- lm(day ~ month + cases_cum + deaths_cum, data = us)
confint(daySTATlm)
summary(daySTATlm)

mse = mean(daySTATlm$residuals^2)
print(paste0("MSE= ", mse))
print(paste0("RMSE= ", RMSE(daySTATlm$residuals)))


par(mfrow = c(2,2)) 
plot(daySTATlm)
```


## Month Model

```{r}
monthSTATlm <- lm(month ~ day + cases_cum + deaths_cum, data = us)
confint(monthSTATlm)
summary(monthSTATlm)

mse = mean(monthSTATlm$residuals^2)
print(paste0("MSE= ", mse))
print(paste0("RMSE= ", RMSE(monthSTATlm$residuals)))


par(mfrow = c(2,2)) 
plot(monthSTATlm)
```


# Linear Models from NY Times COVID-19 GitHub

```{r}
ny_Times_Counties <- read.csv("https://raw.githubusercontent.com/WorldCapital/COVID19-Project/master/COVID-19/nyt-us-counties.csv")
str(ny_Times_Counties)
head(ny_Times_Counties)
```


##Confirmed Model

```{r}
nyTimesCaseslm <- lm(cases ~ fips + deaths, data = ny_Times_Counties)
confint(nyTimesCaseslm)
summary(nyTimesCaseslm)

mse = mean(nyTimesCaseslm$residuals^2)
print(paste0("MSE= ", mse))
print(paste0("RMSE= ", RMSE(nyTimesCaseslm$residuals)))


par(mfrow = c(2,2)) 
plot(nyTimesCaseslm)
```

## Confirmed Model

###Confirmed Model #2 improved R-Squared

```{r}
nyTimesCaseslm2 <- lm(cases ~  deaths, data = ny_Times_Counties)
confint(nyTimesCaseslm2)
summary(nyTimesCaseslm2)

mse = mean(nyTimesCaseslm2$residuals^2)
print(paste0("MSE= ", mse))
print(paste0("RMSE= ", RMSE(nyTimesCaseslm2$residuals)))

par(mfrow = c(2,2)) 
plot(nyTimesCaseslm2)
```

## Deaths Model

```{r}
nyTimesDeathslm <- lm(deaths ~ fips + cases, data = ny_Times_Counties)
confint(nyTimesDeathslm)
summary(nyTimesDeathslm)

mse = mean(nyTimesDeathslm$residuals^2)
print(paste0("MSE= ", mse))
print(paste0("RMSE= ", RMSE(nyTimesDeathslm$residuals)))

par(mfrow = c(2,2)) 
plot(nyTimesDeathslm)
```

## Deaths Model #2

```{r}
nyTimesDeathslm2 <- lm(deaths ~  cases, data = ny_Times_Counties)
confint(nyTimesDeathslm2)
summary(nyTimesDeathslm2)

mse = mean(nyTimesDeathslm2$residuals^2)
print(paste0("MSE= ", mse))
print(paste0("RMSE= ", RMSE(nyTimesDeathslm2$residuals)))

par(mfrow = c(2,2)) 
plot(nyTimesDeathslm2)
```


##FIPS Model

```{r}
nyTimesFIPSlm <- lm(fips ~  cases + deaths, data = ny_Times_Counties)
confint(nyTimesFIPSlm)
summary(nyTimesFIPSlm)

mse = mean(nyTimesFIPSlm$residuals^2)
print(paste0("MSE= ", mse))
print(paste0("RMSE= ", RMSE(nyTimesFIPSlm$residuals)))

par(mfrow = c(2,2)) 
plot(nyTimesFIPSlm)
```

#Non-Linear Regressions 

## Non-Linear Transformation of Confirmed onto Recovered -  STATWORX

```{r}
lm.fit2=lm(cases_cum ∼ deaths_cum + I(deaths_cum^2), data = us) 
summary(lm.fit2)

mse = mean(lm.fit2$residuals^2)
print(paste0("MSE= ", mse))
print(paste0("RMSE= ", RMSE(lm.fit2$residuals)))

par(mfrow = c(2,2))
par()
plot(lm.fit2)
```

##STATWORX NLR #2

```{r}
lm.fit3=lm(deaths_cum ∼ cases_cum + I(cases_cum^2), data = us) 
summary(lm.fit3)

mse = mean(lm.fit3$residuals^2)
print(paste0("MSE= ", mse))
print(paste0("RMSE= ", RMSE(lm.fit3$residuals)))

par(mfrow = c(2,2))
par()
plot(lm.fit3)
```


## Non-Linear Regression of John Hopkins University

```{r}
nonline_fips <- lm(FIPS ~ Confirmed + Deaths + Recovered + Incident_Rate + People_Tested + People_Hospitalized + Mortality_Rate + Testing_Rate + Hospitalization_Rate + I(FIPS^2), data = countyUS)
confint(nonline_fips)
summary(nonline_fips)

mse = mean(nonline_fips$residuals^2)
print(paste0("MSE= ", mse))
print(paste0("RMSE= ", RMSE(nonline_fips$residuals)))

par(mfrow = c(2,2))
plot(nonline_fips)
```


##Non-Linear Model NY Times using Deaths

```{r}
nyTimesNLM <- lm(deaths ~ fips + cases, I(cases^2), data = ny_Times_Counties)
confint(nyTimesNLM)
summary(nyTimesNLM)

mse = mean(nyTimesNLM$residuals^2)
print(paste0("MSE= ", mse))
print(paste0("RMSE= ", RMSE(nyTimesNLM$residuals)))

par(mfrow = c(2,2)) 
plot(nyTimesNLM)
```

#Principal COmponent Analysis (PCA)

```{r, warning=FALSE}
# Principal Component Analysis

countUS <- subset(countyUS, select = -c(Province_State, Last_Update,Country_Region,UID, Lat, Long_, ISO3))

countUS <- na.omit(countUS)

countUS <- transform(countUS, Confirmed = as.numeric(Confirmed), 
          Deaths = as.numeric(Deaths),
          Recovered = as.numeric(Recovered),
          Active = as.numeric(Active),
          FIPS = as.numeric(FIPS),
          People_Tested = as.numeric(People_Tested),
          People_Hospitalized = as.numeric(People_Hospitalized))

apply(countUS, 2, mean)

summary(countUS)



pr.out = prcomp(countUS, scale = FALSE)
names(pr.out)                 


pr.out$center 


pr.out$scale 


pr.out$rotation

dim(pr.out$x)

biplot(pr.out, scale =0)


pr.out$rotation = -pr.out$rotation 
pr.out$x = -pr.out$x 
biplot(pr.out, scale =0) 


pr.out$sdev 

pr.var = pr.out$sdev^2 
pr.var 

pve = pr.var/sum(pr.var) 
pve 

```
pve shows variability >1.